Unsupervised Semantic Mapping for Healthcare Data Storage Schema

### 2021-06-04 ManuScript/IEEEAccess/ManuScript.tex
fixed class labels

This annotated data was then processed, to replace all related entries with ``0.5" and all``-" with ``0", while the values for similar at ``1" and ``0" for dissimilar were kept the same. This conversion was then used to produce a consolidated dataset of 40,698 attribute pairs, using mode scores of all annotators, for each cell. We also tested average scores between the annotators, but that would produce scores between ``0", ``0.5", and ``1", greatly increasing the number of classes for classification. Hence the maximum agreement between the annotators, maintains the final label values within these three classes, which become easier to evaluate. Additionally, the original dataset and its mode consolidated form is biased in favour of class "0", since most attribute pairs are not related to each other.
This dataset is then split into development and testing partitions with a ratio of 70:30. The development partition is used for threshold selection, based on the best MCC score for identifying class ``equal", followed by best scores for class ``related" and finally best of class "unrelated". The optimal threshold thus achieved is used to classify the instances of the test dataset, which is finally evaluated on its MCC and F-1 score.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the Matthews Correlation Coefficient (MCC), is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem (class 0, class 0.5, and class 1), we have to test various threshold scores for separating the instances between 0 and 0.5 ($t_1$), and then 0.5 and 1($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of class 1 (similar), followed by class 0.5 (related), and finally class 0 (unrelated).
With a step size of 0.05 ($step_size$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step_size$, we move the thresholds until $t_2$ reaches 1, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches 0.95 and $t_2$ reaches 1, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy", ``Word2Vec", and 10 BERT based models.

Threshold values for Word2Vec are placed at the lower end of the spectrum, indicating a very large number of instances are classified as similar (above similarity score of 0.1), while a small number of instances (with similarity score 0.05) are classified as dissimilar. Similarly, the class related lies within the similarity threshold of 0.05 similarity points. It can be observed that the threshold for selecting the related class, is within 0.05 points, in all except one case (bert-large-nli-mean-tokens where the difference is 0.15 points). Five BERT based models, trained on the STSb dataset, all have minimum threshold values of 0.85 and maximum of 0.9, while the 4 remaining BERT models, lie between 0.9 and 0.95. These results show a general trend, of how the cosine similarity varies/maintains itself, against embedded vectors generated from various pre-trained models. In absolute terms, however these threshold values provide the mechanism for classifying the test dataset, which is evaluated for performance, in the next subsection.

\subsection{Model Evaluation}
On unseen test dataset, with thresholds selected in the previous step and the 12 models, we measured the performance score, using one vs all binarization of the multi-classes. For class 1 (``similar")

@@

		\multirow{3}{*}{FUZZY\_MATCH}                         & 0.0                     & 0.5,1                   & 0.98                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.37                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.98                                   & 0.20                                    & 0.59                                 & 0.29                              & 0.33                              \\ \hline
		\multirow{3}{*}{Word2Vec}                             & 0.0                     & 0.5,1                   & 0.24                                   & 0.99                                    & 0.24                                 & 0.38                              & 0.01                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.15                                    & 0.10                                 & 0.12                              & 0.12                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.35                                   & 0.01                                    & \textbf{0.84}                        & 0.01                              & 0.03                              \\ \hline
		\multirow{3}{*}{bert-base-nli-stsb-mean-tokens}       & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.37                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.22                                    & 0.50                                 & 0.30                              & 0.32                              \\ \hline
		\multirow{3}{*}{bert-large-nli-stsb-mean-tokens}      & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 1.00                                 & 1.00                              & 0.51                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.41                                    & 0.50                                 & 0.45                              & 0.45                              \\ \hline
		\multirow{3}{*}{roberta-base-nli-stsb-mean-tokens}    & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 1.00                              & 0.40                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.26                                    & 0.50                                 & 0.34                              & 0.36                              \\ \hline
		\multirow{3}{*}{roberta-large-nli-stsb-mean-tokens}   & 0.0                     & 0.5,1                   & 1.00                                   & 1.00                                    & 1.00                                 & 1.00                              & 0.61                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & \textbf{1.00}                          & \textbf{0.59}                           & 0.50                                 & \textbf{0.54}                     & \textbf{0.54}                     \\ \hline
		\multirow{3}{*}{distilbert-base-nli-stsb-mean-tokens} & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.38                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.23                                    & 0.50                                 & 0.32                              & 0.33                              \\ \hline
		\multirow{3}{*}{bert-base-nli-mean-tokens}            & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.35                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.20                                    & 0.50                                 & 0.29                              & 0.31                              \\ \hline
		\multirow{3}{*}{bert-large-nli-mean-tokens}           & 0.0                     & 0.5,1                   & 0.28                                   & 1.00                                    & 0.27                                 & 0.43                              & 0.05                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 0.99                                   & 0.34                                    & 0.90                                 & 0.49                              & 0.55                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.57                                   & 0.00                                    & 0.50                                 & 0.01                              & 0.01                              \\ \hline
		\multirow{3}{*}{roberta-base-nli-mean-tokens}         & 0.0                     & 0.5,1                   & 0.96                                   & 1.00                                    & 0.97                                 & 0.98                              & 0.23                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.96                                   & 0.08                                    & 0.50                                 & 0.13                              & 0.19                              \\ \hline
		\multirow{3}{*}{roberta-large-nli-mean-tokens}        & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.39                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.20                                    & 0.50                                 & 0.28                              & 0.31                              \\ \hline
		\multirow{3}{*}{distilbert-base-nli-mean-tokens}      & 0.0                     & 0.5,1                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.36                              \\ \cline{2-8} 
		& 0.5                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& 1                       & 0.0,0.5                 & 0.99                                   & 0.21                                    & 0.50                                 & 0.30                              & 0.32                              \\ \hline

>>>>

This annotated data was then processed, to replace all related entries with ``0.5" ($class_{0.5}$) and all``-" with ``0" ($class_{0}$), while the values for similar at ``1" ($class_{1}$) and ``0" ($class_{0}$) for dissimilar were kept the same. This conversion was then used to produce a consolidated dataset of 40,698 attribute pairs, using mode scores of all annotators, for each cell. We also tested average scores between the annotators, but that would produce scores between ``0", ``0.5", and ``1", greatly increasing the number of classes for classification. Hence the maximum agreement between the annotators, maintains the final label values within these three classes, which become easier to evaluate. Additionally, the original dataset and its mode consolidated form is biased in favour of class "0", since most attribute pairs are not related to each other.
This dataset is then split into development and testing partitions with a ratio of 70:30. The development partition is used for threshold selection, based on the best MCC score for identifying class ``equal", followed by best scores for class ``related" and finally best of class "unrelated". The optimal threshold thus achieved is used to classify the instances of the test dataset, which is finally evaluated on its MCC and $F_1$ measure.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the Matthews Correlation Coefficient (MCC), is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem ($class_0$, $class_{0.5}$, and $class_1$), we have to test various threshold scores for separating the instances between $class_0$ and $class_{0.5}$ ($t_1$), and then $class_{0.5}$ and $class_1$ ($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of $class_1$ (similar), followed by $class_{0.5}$ (related), and finally $class_0$ (unrelated).
With a step size of \textit{0.05} ($step$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step$, we move the thresholds until $t_2$ reaches \textit{1.0}, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches \textit{0.95} and $t_2$ reaches \textit{1.0}, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy", ``Word2Vec", and 10 BERT based models.

Threshold values for Word2Vec are placed at the lower end of the spectrum, indicating a very large number of instances are classified as similar (above similarity score of \textit{0.1}), while a small number of instances (with similarity score \textit{0.05}) are classified as dissimilar. Similarly, the $class_{0.5}$ lies within the similarity threshold of \textit{0.05} similarity points. It can be observed that the threshold for selecting the related class, is within \textit{0.05} points, in all except one case (bert-large-nli-mean-tokens where the difference is \textit{0.15} points). Five BERT based models, trained on the STSb dataset, all have minimum threshold values of \textit{0.85} and maximum of \textit{0.9}, while the 4 remaining BERT models, lie between \textit{0.9} and \textit{0.95}. These results show a general trend, of how the cosine similarity varies/maintains itself, against embedded vectors generated from various pre-trained models. In absolute terms, however these threshold values provide the mechanism for classifying the test dataset, which is evaluated for performance, in the next subsection.

\subsection{Model Evaluation}
On unseen test dataset, with thresholds selected in the previous step and the 12 models, we measured the performance score, using one vs all binarization of the multi-classes. As evident in Table \ref{tab:evaluations}, very high values of accuracy are visible across all models, with all three positive classes. In all, except the case of Word2Vec, Precision and recall, also show values close to \textit{1.0}. However, these measures are very misleading, since the dataset is greatly biased in favour of $class_0$. In terms of $F_1$ measure $class_{0.5}$ shows the worst possible results, independently, with all except Word2Vec and bert-large-nli-mean-tokens, having a score of \textit{0.0}. Bert-large-nli-mean-tokens, provides the best $F_1$ measure at \textit{0.49}. These metrics are thus not useful to gauge the performance of the evaluated models. Instead, focusing on the MCC score, provides a good picture of the model performance, for individual classes, when all other instances are negative.

@@

		\multirow{3}{*}{FUZZY\_MATCH}                         & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.98                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.37                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.98                                   & 0.20                                    & 0.59                                 & 0.29                              & 0.33                              \\ \hline
		\multirow{3}{*}{Word2Vec}                             & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.24                                   & 0.99                                    & 0.24                                 & 0.38                              & 0.01                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.15                                    & 0.10                                 & 0.12                              & 0.12                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.35                                   & 0.01                                    & \textbf{0.84}                        & 0.01                              & 0.03                              \\ \hline
		\multirow{3}{*}{bert-base-nli-stsb-mean-tokens}       & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.37                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.99                                   & 0.22                                    & 0.50                                 & 0.30                              & 0.32                              \\ \hline
		\multirow{3}{*}{bert-large-nli-stsb-mean-tokens}      & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 1.00                                 & 1.00                              & 0.51                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.99                                   & 0.41                                    & 0.50                                 & 0.45                              & 0.45                              \\ \hline
		\multirow{3}{*}{roberta-base-nli-stsb-mean-tokens}    & $class_0$                    & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 1.00                              & 0.40                              \\ \cline{2-8} 
		& $class_{0.5}$                     & 0.0,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                & 0.99                                   & 0.26                                    & 0.50                                 & 0.34                              & 0.36                              \\ \hline
		\multirow{3}{*}{roberta-large-nli-stsb-mean-tokens}   & $class_0$                     & $class_{0.5}$,$class_1$                   & 1.00                                   & 1.00                                    & 1.00                                 & 1.00                              & 0.61                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & 0.0,$class_{0.5}$                & \textbf{1.00}                          & \textbf{0.59}                           & 0.50                                 & \textbf{0.54}                     & \textbf{0.54}                     \\ \hline
		\multirow{3}{*}{distilbert-base-nli-stsb-mean-tokens} & 0.0                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.38                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.99                                   & 0.23                                    & 0.50                                 & 0.32                              & 0.33                              \\ \hline
		\multirow{3}{*}{bert-base-nli-mean-tokens}            & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.35                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.99                                   & 0.20                                    & 0.50                                 & 0.29                              & 0.31                              \\ \hline
		\multirow{3}{*}{bert-large-nli-mean-tokens}           & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.28                                   & 1.00                                    & 0.27                                 & 0.43                              & 0.05                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 0.99                                   & 0.34                                    & 0.90                                 & 0.49                              & 0.55                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.57                                   & 0.00                                    & 0.50                                 & 0.01                              & 0.01                              \\ \hline
		\multirow{3}{*}{roberta-base-nli-mean-tokens}         & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.96                                   & 1.00                                    & 0.97                                 & 0.98                              & 0.23                              \\ \cline{2-8} 
		& $class_{0.5}$                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                 & 0.96                                   & 0.08                                    & 0.50                                 & 0.13                              & 0.19                              \\ \hline
		\multirow{3}{*}{roberta-large-nli-mean-tokens}        & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.39                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$                & 0.99                                   & 0.20                                    & 0.50                                 & 0.28                              & 0.31                              \\ \hline
		\multirow{3}{*}{distilbert-base-nli-mean-tokens}      & $class_0$                     & $class_{0.5}$,$class_1$                   & 0.99                                   & 1.00                                    & 0.99                                 & 0.99                              & 0.36                              \\ \cline{2-8} 
		& $class_{0.5}$                     & $class_0$,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 
		& $class_1$                       & $class_0$,$class_{0.5}$           & 0.99                                   & 0.21                                    & 0.50                                 & 0.30                              & 0.32                              \\ \hline
